---
title: "High School Recruitment Strategies"
author: "Ryder Cobean; https://github.com/ryderdavid/nyc_school_recruitment" 
output: 
  output: html_document
---

<style>
.leaflet {
    margin: auto;
}
</style>



This report attempts to answer a number of questions for a School in the South Bronx, New York City. These analyses were done based off of data from that school. All data had no individual student identifiers. 

## Bus Advertisement Targeting

### Marketing for English Language Learner Households

One angle for targetting recruitment efforts is examining the distribution of "limited English proficiency" households in New York. The [American Community Survey](https://www.census.gov/programs-surveys/acs/news/data-releases.html?# "American Community Survey 2017 Data Releases") tabulates limited English proficiency households by census tract, which are defined as households that do not solely speak English, in which at least one member is reported by the respondent to speak English "not very well."
```{r lep_C16002_setup}

# GET 2017 5YR ACS HOUSEHOLD LANGUAGE SPOKEN BY LIMITED ENGLISH SPEAKING STATUS
options(tigris_use_cache = TRUE)
lep_hh_by_tract_C16002 <- 
  get_acs(geography = "tract", year = 2017, state = "NY", 
          county = c("New York", "Bronx", "Richmond", "Queens", "Kings"),
          output = "wide",
          geometry = T,
          variables = c(total_hh = "C16002_001E", 
                        spanish_lep = "C16002_004E",
                        indeur_lep = "C16002_007E",
                        asnpacisl_lep = "C16002_010E",
                        otherlang_lep = "C16002_013E",
                        total_pop = "B01003_001E")) %>% 
  st_transform(crs = wgs84_crs) %>% 
  st_erase(., nyc_water)

# get County name from longform tract name in NAME, get boro name from county names
# Don't need this anymore since adding boro
lep_hh_by_tract_C16002 %<>%
  mutate(county_name = str_extract(NAME, "((?<=\\d, )[:graph:]+).+(?=\\sCounty)")) 

# break out the data from the sf object and summarize across rows for each
# language's limited english proficiency respondents. Make sure to not introduce
# NaN's by dividing by zero!
total_lep_by_tract <- lep_hh_by_tract_C16002 %>% as_tibble() %>% 
  
  # Filter out first percentile by pop tracts (often parks and cemeteries)
  filter(total_pop >= quantile(total_pop, 0.1)) %>% 
  
  # get a variable summing LEP counts of each language group as total_lep
  transmute(total_lep = dplyr::select(., spanish_lep, indeur_lep, 
                                      asnpacisl_lep, otherlang_lep) %>% rowSums(),
            
            # and one for percentage. If total_lep is 0, hard enter 0% for pct
            pct_lep = ifelse(total_lep == 0, 0, 100 * total_lep / total_hh), 
            
            #keep GEOID variable
            GEOID = GEOID)

# merge the data back into the geometry
lep_hh_by_tract_C16002 <- inner_join(lep_hh_by_tract_C16002, 
                                     total_lep_by_tract, 
                                     by = c("GEOID" = "GEOID"))

# prep for plotting: add school districts for each tract.
lep_hh_by_tract_C16002 <- st_join(lep_hh_by_tract_C16002, nyc_sds)

# filter to just Manhattan and Bronx
lep_hh_by_tract_C16002_manbx <- filter(lep_hh_by_tract_C16002,
                                       county_name %in% c("New York", "Bronx"))


```

The below interactive map shows the percentage of limited English proficiency households per census tract in NYC, with school district boundaries overlaid. __Note that this calculation has taken out of consideration those tracts with extremely low (first percentile and below) total tract populations.__ While some of those tracts have high proportions of limited English proficiency households, the total number of people living in them is extremely small, and most of them represent land that is largely a park or cemetery.

```{r LEP_household_proportions_map}

# Plot map of LEP household proportions by tract
tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  tm_shape(nyc_sds, point.per = "feature", bbox = man_bx_bbox) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  tm_shape(lep_hh_by_tract_C16002) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep"))



```



```{r bus_shelters_lep_setup, cache=T, dependson=c("setup", "functions"), include=FALSE}
pct <- 0.80
high_lep_hh_manbx <- 
  lep_hh_by_tract_C16002_manbx %>% 
  filter(pct_lep >= quantile(.$pct_lep, pct))


# Get all bus shelters in NYC that are in pctile-th-percentile tracts by proportion
# of LEP households


# 
# high_lep_bus_shelters_manbx <- 
#   nyc_bus_shelters[high_lep_hh_manbx, ]

high_lep_bus_shelters_manbx <- st_intersection(nyc_bus_shelters, high_lep_hh_manbx)

# Prep a version of the dataframe of 
shelter_table <- high_lep_bus_shelters_manbx %>% as_tibble() %>% 
  transmute(`Shelter ID` = shelter_id, Coordinates = coords,
            Location = paste(location, " at ", str_to_title(at_between), sep=""),
            `Census Tract` = GEOID,
            `LEP Households` = total_lep) %>% 
  arrange(desc(`LEP Households`))

```

### Bus Shelters in High-LEP Census Tracts
The following `r as.character(nrow(high_lep_bus_shelters_manbx))` bus shelters are located in the tracts in Manhattan and the Bronx with the highest (`r as.character(pct * 100)`th percentile) proportions of limited english proficiency. Again, census tracts with extremely low population (often parks, cemeteries and other such areas) have been removed prior to calculation.


```{r bus_shelters_lep_plot}
tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  tm_shape(nyc_sds, point.per = "feature", 
           bbox = st_bbox(filter(high_lep_bus_shelters_manbx, boro_name == "Bronx"))) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  tm_shape(high_lep_hh_manbx) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep")) +
  tm_shape(high_lep_bus_shelters_manbx) +
  tm_dots(col = "#42f4e2",
          id = "shelter_id",
          popup.vars = c("Location" = "location",
                         "Coordinates" = "coords",
                         "Nearby LEP Households" = "total_lep",
                         "% LEP households in Tract" = "pct_lep"))

```

\  
\  


```{r}
kable(shelter_table, caption = "Bus shelters in proportionally high-LEP household tracts") %>%
  kable_styling(bootstrap_options = "striped") %>% 
  scroll_box(height = "300px")

```
\  

The entire table can be copied and pasted into a spreadsheet and will retain its formatting.
\  

```{r bus_routes}

st_intersects_tally_attr <-
  function(x = NULL,
           y = NULL,
           attribute = NULL) {
    require(sf)
    
    # a <- attribute
    a <- attribute
    
    
    # a list of x's length with each record containing a vector of indices of y's
    # records that intersect with that record in x
    xy <- st_intersects(x, y)
    
    
    attr_n <- c() # empty vector to fill
    for (r in 1:length(xy)) {
      n <- 0
      
      if (length(xy[[r]]) == 0) {
        n <- 0
        
      } else {
        for (i in 1:length(xy[[r]])) {
          n <- n + st_drop_geometry(y)[i, a]
          
        }
        
      }
      
      attr_n <- c(attr_n, n)
      
    }
    
    
    return(x %>% bind_cols((!!a) := attr_n))
    
  }

# Create a subset of buses that just run in Manhattan and the Bronx
manbx_bus_sgbp <- st_intersects(nyc_bus_routes, c(manhattan_sf, bronx_sf))
manbx_bus_routes <- nyc_bus_routes[lengths(manbx_bus_sgbp) > 0, ]
rm(manbx_bus_sgbp)  # get rid of helper variable

# Tally the value "total_lep" for its intersects with bus routes passing through
# tracts with the attribute
manbx_bus_routes_with_lep <- st_intersects_tally_attr(manbx_bus_routes,
                                                      lep_hh_by_tract_C16002_manbx,
                                                      "total_lep")



manbx_bus_routes_with_lep %<>% 
  top_n(10, total_lep) %>% 
  arrange(desc(total_lep)) %>% 
  mutate(route_id = fct_drop(route_id))


```

### Bus Routes Along High-LEP Areas

In addition to advertising on the bus stops located within the highest-proportion LEP census tracts in New York, placing advertisements for the school on the bus lines passing near the most LEP households is another strategy. Below are the top `r nrow(manbx_bus_routes_with_lep)` routes in Manhattan and the Bronx, by the number of Low English Proficiency households they pass by (total LEP household counts of the census tracts they pass through):
  
  
```{r bus_route_lep_plot}




tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  
  #  NYC School Districts Base - SD Borders
  tm_shape(nyc_sds, point.per = "feature") + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  
  #  Polygons of high proportion Low English Proficiency Tracts  
  tm_shape(high_lep_hh_manbx) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          # title = "% LEP Households",
          id = "NAME",
          legend.show = F,
          showNA = F,
          alpha = 0.5,
          popup.vars = c("Tract ID: "="GEOID", 
                         "Name: "="NAME", 
                         "School District: "="school_dist",
                         "Total Households: "="total_hh",
                         "% LEP Households: "="pct_lep")) +
  
  #  Top n bus routes that pass through the census tracts with the most LEP
  #  households
  tm_shape(manbx_bus_routes_with_lep) + 
  tm_lines(lwd = 3,
           id = "route_id",
           col = "route_id",
           palette = "Accent",
           title.col = "Route",
           alpha = 1,
           popup.vars = c("Route ID: " = "route_id",
                          "Name: " = "route_long",
                          "# LEP Households: " = "total_lep")) +
  
  #  Focus view to the center of the Bronx, set zoom to fill view with Bronx
  tm_view(set.view=c(-73.88251, 40.83489, 12),
          set.zoom.limits = c(11, 15))

```
\  
\  
```{r bus_route_lep_table}

manbx_bus_routes_with_lep_kable <- manbx_bus_routes_with_lep %>% 
  dplyr::select(`Route ID`=route_id, 
                `Route Name`=route_long,
                `LEP Households`=total_lep) %>% 
  st_drop_geometry()

kable(manbx_bus_routes_with_lep_kable, 
      caption = paste("Top ", nrow(manbx_bus_routes_with_lep_kable), " bus routes by LEP Household Passbys, MAN/BX")) %>%
  kable_styling(bootstrap_options = "striped")


```

\  

## Applications Needed to Reach a Target Enrollment

Though data has not been collected over sufficient years to make an authoritative prediction about the number of applications needed to recruit to a target enrollment, a rough estimate can be derived from the applications to registrations rate in the years with recorded registrations. 

The below table outlines the current application to registration performance:
  
```{r app_to_reg}

recruitment_data %>% 
  st_drop_geometry() %>% 
  count(year, registered) %>% 
  complete(year, registered, fill = list(n = 0)) %>% 
  spread(registered, n) %>% 
  transmute(Year = year,
            `Not Registered` = `FALSE`,
            `Registered` = `TRUE`,
            `Total Applications` = `Not Registered` + `Registered`) %>% 
  kable(., caption = "Total Applications and Registrations, 2017-2019 School Years") %>% 
  kable_styling(bootstrap_options = "striped")



```


Here, the below outlines the estimated applications that might be necessary to hit a target recruitment. This could be used as a hypothesis to test in coming years as further data comes in. A better approach could be looking at application to registration performance in top recruiting neighborhoods to see how they perform over time if recruitment efforts are focused in these neighborhoods over others, or to test whether increased or diversified efforts in underperforming neighborhoods may result in better performance in those areas. 

```{r applications_needed_to_reach_target}

recruitment_data %>% 
  filter(year %in% c(2017, 2018)) %>% 
  st_drop_geometry() %>% 
  count(registered) %>% 
  spread(registered, n) %>% 
  transmute(Applied = `TRUE` + `FALSE`,
            Registered = `TRUE`,
            `Registration Rate` = Registered / Applied,
            `To Reach 1200` = 1200 %/% `Registration Rate`) %>% 
  kable(., caption = "Applications needed to reach target enrollment, based on 2017-2018 recruitment performance") %>% 
  kable_styling(bootstrap_options = "striped")

```

\  
\  
