---
title: "High School Recruitment Report"
author: "Ryder Cobean"
output: 
  output: html_document
---

<style>
.leaflet {
    margin: auto;
}
</style>



This report attempts to answer a number of questions for a School in the South Bronx, New York City. These analyses were done based off of data from that school. All data had no individual student identifiers. 


### Application Data {.tabset}

Below are the distributions of applications received from the `r applications_by_tract_yearly_totals$year %>% min()`-`r (applications_by_tract_yearly_totals$year %>% min() + 1)` to the `r applications_by_tract_yearly_totals$year %>% max()`-`r applications_by_tract_yearly_totals$year %>% max() + 1` school years, arranged by NYC Census Tract. In total, `r applications_by_tract_yearly_totals$n %>% sum()` applications were received from `r applications_by_tract_yearly_totals$tractname %>% unique() %>% length()` tracts.


```{r applications_plot}

tmap_mode("view")

# The below pre-render choropleth maps for 2017, 2018, and 2019's total applications


plot_app_tract_totals <- function(yr) {
  
  plt <- tm_tiles(providers$CartoDB.PositronNoLabels) + 
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(applications_by_tract_yearly_totals %>% 
               filter(year == yr)) + 
      tm_borders(alpha = 0.2) +
      tm_fill(col = "n",
            id="tractname",
            breaks = apps_jenks_breaks,
            title = paste(yr, " Apps", sep = ""),
            popup.vars = c("Applications"="n",
                           "% of Year's Applications"="pct_yr")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
            set.zoom.limits = c(11,14)) + 
    tm_layout(outer.margins = 20)  
  
  tmap_mode("view")
  return (plt)
  
}




```
  
#### 2017 Applications  
  
```{r}
plot_app_tract_totals(2017)
```

#### 2018 Applications  
  
```{r}
plot_app_tract_totals(2018)
```

#### 2019 Applications  
  
```{r}
plot_app_tract_totals(2019)
```

### Registration Data {.tabset}

Below are the distributions of applications that resulted in the student being accepted and registered. The registration data currently tracks the `r min(registrations_by_tract_yearly_totals$year)`-`r min(registrations_by_tract_yearly_totals$year) + 1` to the `r max(registrations_by_tract_yearly_totals$year)`-`r max(registrations_by_tract_yearly_totals$year) + 1` school years, with `r sum(registrations_by_tract_yearly_totals$n)` registrations in total occurring from `r registrations_by_tract_yearly_totals$tractname %>% unique() %>% length()` tracts in New York City.

```{r registrations_plot}

tmap_mode("view")

plot_reg_tract_totals <- function(yr) {
  
  plt <- tm_tiles(providers$CartoDB.PositronNoLabels) +
    tm_shape(nyc_sds, point.per = "feature") +
    tm_borders(alpha = 0.5) +
    tm_text(text = "school_dist") +
    tm_shape(registrations_by_tract_yearly_totals %>%
             filter(year == yr)) +
    tm_borders(alpha = 0.2) +

    tm_fill(col = "n",
            palette = "YlOrBr",
            id="tractname",
            breaks = regs_jenks_breaks,
            title = paste(yr, " Regs", sep = ""),
            popup.vars = c("Registrations"="n",
                           "% of Year's Registrations"="pct_yr")) +
    tm_view(set.view = c(-73.86652, 40.85263, 11),
          set.zoom.limits = c(11,14))
  
  return(plt)
  
}


```
  
#### 2017 Registrations
```{r}
plot_reg_tract_totals(2017)
```

#### 2018 Registrations
```{r}
plot_reg_tract_totals(2018)
```


  
### Application to Registration Performance {.tabset}

```{r recruitment_performance_setup, include=F}

tracts_by_recruitment_performance <- 
  recruitment_data %>% 
  count(year, GEOID, registered) %>% 
  spread(key = registered, value = n, fill = 0) %>% 
  complete(year, GEOID, fill = list(`TRUE` = 0, `FALSE` = 0)) %>% 
  transmute(Year = year,
            GEOID = GEOID,
            Applied = `TRUE` + `FALSE`,
            `Not Registered` = `FALSE`,
            Registered = `TRUE`,
            `Recruitment Rate` = ifelse(Applied == 0, 0, Registered / Applied))


tracts_by_recruitment_performance <- 
  recruitment_data %>% 
  count(year, GEOID, registered) %>% 
  spread(key = registered, value = n, fill = 0) %>% 
  transmute(Year = year,
            GEOID = GEOID,
            Applied = `TRUE` + `FALSE`,
            `Not Registered` = `FALSE`,
            Registered = `TRUE`,
            `Recruitment Rate` = ifelse(Applied == 0, 0, Registered / Applied)) %>% 
  st_drop_geometry() %>% 
  inner_join(., nyc_tracts, by = c("GEOID" = "GEOID")) %>% 
  st_as_sf(.)

# Applications to Registrations Rate Map:
apps_to_regs_map <- function(years = seq(0,10000), 
                             pal = "YlOrBr") {
  
  data <- tracts_by_recruitment_performance
  years <- years
  pal <- pal

  data %<>%
    filter(Year %in% years)

  map_plot <-
    tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) +
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(data) +
      tm_borders(alpha = 0.2) +
      tm_fill(col = "Recruitment Rate",
              palette = pal,
              id = "tractname",
              title =

                # logic to handle whether years renders as
                # YYYY-YY or just YYYY depending on whether
                # multiple years are called in args
                if (length(years) > 1) {
                  paste("Regs:Apps, ", "\n", min(years), "-",
                            substr(toString(max(years)), 3, 4),
                            sep = "")
                } else {
                  paste("Apps:Regs, ", "\n",
                        years,
                        sep = "")
                },

              popup.vars =
                c("Tract ID #"="GEOID",
                  "Applied",
                  "Registered",
                  "Not Registered",
                  "Recruitment Rate")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
              set.zoom.limits = c(11,14)) +
      tm_layout(outer.margins = 20)
  
  return(map_plot)
  
}

```


Below are a number of views of the recruitment performance, by different geographies.

#### By School District, All Years

Below is the comparative performance of recruitment in each of New York's `r nyc_sds %>% length()` school district boundaries, showing the number of total applications received and the number of registrations completed from those applications, per year, for the `r min(recruitment_by_sd_tidy$year)`-`r min(recruitment_by_sd_tidy$year) + 1` to the `r max(recruitment_by_sd_tidy$year)`-`r max(recruitment_by_sd_tidy$year) + 1` school years.


```{r app_reg_performance}


ggplot(recruitment_by_sd_tidy %>% 
         mutate(school_dist = as.factor(school_dist))) + 
  geom_bar(aes(x = school_dist, y = n, fill = registered), stat = "identity") +
  ggtitle("Recruitment Performance by Applicant's School District", 
          subtitle = "School Years Beginning 2017 through 2019") +
  theme(plot.title      = element_text(hjust = 0.5),
        plot.subtitle   = element_text(hjust = 0.5),
        legend.position = "bottom") +
  xlab("School District") + 
  ylab("Count") +
  facet_grid(year ~ .)




```


#### By Census Tracts, 2017

Below are census tracts depicting rate of registration (total registrations / total applications, by applications originating in the census tract.)
```{r}
apps_to_regs_map(2017)
```

#### By Census Tracts, 2018

Below are census tracts depicting rate of registration (total registrations / total applications, by applications originating in the census tract.)
```{r}
apps_to_regs_map(2018, "BuGn")
```



\  
\  
