---
title: "HUM II Recruitment"
author: "Ryder Cobean"
output: 
  html_document:
    code_folding: hide
---

This report attempts to answer a number of questions for Humanities II Charter School in The Bronx, New York City.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', message = F, warning = F)
```

```{r, setup, results='hide'}

check.packages <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# need local package gdal, units (udunits on arch), v8-3.14 (source on AUR), gcc-fortran, unixodbc

packages <- c('here', 'sf', 'raster', 'devtools', 'acs', 'tidycensus', 'tidyverse', 'tigris', 'sp', 
              'tmap', 'tmaptools', 'readxl', 'ggplot2', 'rgdal', 'spdplyr', 'RColorBrewer', 
              'viridis', 'viridisLite', 'rstudioapi', 'magrittr', 'getPass', "kableExtra", 'rmarkdown')

check.packages(packages)


if(!dir.exists("data")) {
  dir.create("data/downloads", recursive = T)
} else if (!dir.exists("data/downloads")) {
  dir.create("data/downloads")
}


if(Sys.getenv("CENSUS_API_KEY") == "") {
  
  census_api_key(getPass("Enter U.S. census API key to install: "), install = T, overwrite = T)
  readRenviron("~/.Renviron")
  
}


##### SET A REFERENCE VARIABLE HERE: CRS FOR ALL LAYERED PLOTS ##### use this
# CRS to transform projeections of any other layer being used when necessary

wgs84_crs <- "+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"




# Download shapefiles for NYC area ZCTAs to render as background layer
options(tigris_use_cache = T)
nyc_area_zips <- 
  zctas(cb = T, starts_with = c('070','071','072','073','074',
                                '075','076','100','101','102',
                                '103','104','105','106','107',
                                '108','109','110','111','112',
                                '113','114','115','116')) %>% 
  as("sf") %>%
  st_transform(crs = wgs84_crs)


# get a simple block of NJ coastline to add to baselayers of maps
nj_land <- nyc_area_zips %>% filter(str_detect(GEOID10, "^07")) %>% st_union
  

# Filter just the ZCTAs we're interested in - Man/Bx
man_bx_zips <- nyc_area_zips %>% filter(str_detect(GEOID10, "^100|^101|^102|^103|^104"))


# Create a "fixed" shapefile of NYC School Districts - merging all of SD 10
# which is split over Bronx / Manhattan boundary. School District Shapefile
# from:
# https://www1.nyc.gov/assets/planning/download/zip/data-maps/open-data/nysd_18d.zip
# 
# download.file("https://www1.nyc.gov/assets/planning/download/zip/data-maps/open-data/nysd_18d.zip", 
#   destfile = "data/nysd_18d.zip")
# unzip("nysd_18d.zip", exdir = "data/nysd_18d", overwrite = T)

# NYC's SD geojson file only pulls in data as characters, which leads sf to read
# them as factors which breaks all attempts at analysis. forcing NOT to factors
# brings everything in as characters, which lets us cast each attribute to
# numeric
nyc_sds <- st_read("https://data.cityofnewyork.us/resource/cuae-wd7h.geojson", 
                   stringsAsFactors = F) %>% 
  st_transform(crs = wgs84_crs) %>% 
  mutate(school_dist = as.numeric(school_dist), 
         shape_area = as.numeric(shape_area),
         shape_leng = as.numeric(shape_leng)) %>% 
  group_by(school_dist) %>% 
  summarize()




districts_demo_snapshot <- 
  read_csv("https://data.cityofnewyork.us/resource/dndd-j759.csv")



names(districts_demo_snapshot) %<>% 
  str_to_lower() %>% 
  str_replace_all("\\s", "_") %>% 
  str_replace_all("#", "num") %>% 
  str_replace_all("%", "pct") %>% 
  str_replace_all("\\(|\\)", "") %>% 
  str_replace_all("\\&", "") %>% 
  str_replace_all("_{2,}", "_")

districts_demo_snapshot <- districts_demo_snapshot %>% 
  rename(num_swd = students_with_disabilities_1, 
         pct_swd = students_with_disabilities_2,
         num_ell = english_language_learners_1, 
         pct_ell = english_language_learners_2, 
         num_pov = poverty_1, 
         pct_pov = poverty_2)



districts_demo_snapshot_1718 <- districts_demo_snapshot %>% 
  filter(year == "2017-18")

districts_demo_snapshot_1718 <- 
  inner_join(nyc_sds, districts_demo_snapshot, 
             by = c("school_dist" = "administrative_district"))





schools_demo_snapshot <- 
  read_csv("https://data.cityofnewyork.us/api/views/s52a-8aq6/rows.csv?accessType=DOWNLOAD&bom=true&format=true")

names(schools_demo_snapshot) %<>% 
  str_to_lower() %>% 
  str_replace_all("\\s", "_") %>% 
  str_replace_all("#", "num") %>% 
  str_replace_all("%", "pct") %>% 
  str_replace_all("\\(|\\)", "") %>% 
  str_replace_all("\\&", "") %>% 
  str_replace_all("_{2,}", "_")

schools_demo_snapshot <- schools_demo_snapshot %>% 
  rename(num_swd = num_students_with_disabilities, pct_swd = pct_students_with_disabilities,
         num_ell = num_english_language_learners, pct_ell = pct_english_language_learners,
         num_pov = num_poverty, pct_pov = pct_poverty)


# Shapefile of public school points - This is an ESRI shape file of school point
# locations based on the official address.  It includes some additional basic
# and pertinent information needed to link to other data sources. It also
# includes some basic school information such as Name, Address, Principal, and
# Principal’s contact information.

download.file(url = "https://data.cityofnewyork.us/download/jfju-ynrr/application%2Fzip", destfile = "data/downloads/school_points.zip")
unzip("data/downloads/school_points.zip", exdir = "data/school_points", overwrite = T)

school_points <- st_read("data/school_points/Public_Schools_Points_2011-2012A.shp")

names(school_points) %<>% 
  str_to_lower()


# clean up trailing unicode characters in ats_code
school_points$ats_code %<>% str_sub(1,6)

school_points <- st_transform(school_points, crs = wgs84_crs)

# add point geometry to school demographic data
schools_demo_snapshot <- inner_join(schools_demo_snapshot, school_points, by = c("dbn" = "ats_code")) %>% st_as_sf()

# filter out old years of schools demo snapshot
schools_demo_snapshot_1718 <- schools_demo_snapshot %>% filter(year == "2017-18")


# Get census tract shapefiles

# NYC tracts clipped to shoreline available here: 
# https://www1.nyc.gov/site/planning/data-maps/open-data/districts-download-metadata.page
nyc_tracts <- 
  st_read("http://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/nyct2010/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=geojson")


nyc_tracts <- nyc_tracts %>% st_transform(crs = wgs84_crs)


# The below sets up the necessary shapefile and function to clip any shapefile
# to the NYC shoreline (for example, the US Census Tract boundaries, which
# extend into the Hudson and East Rivers). Passing your shapefile and nyc_water
# to st_erase will clip them to the shoreline. Curiously, 
st_erase <- function(x, y) {
  st_difference(x, st_union(st_combine(y)))
}

bx_water <- tigris::area_water("NY", county = "Bronx", class = "sf")
bk_water <- tigris::area_water("NY", county = "Kings", class = "sf")
qn_water <- tigris::area_water("NY", county = "Queens", class = "sf")
st_water <- tigris::area_water("NY", county = "Richmond", class = "sf")
ny_water <- tigris::area_water("NY", county = "New York", year = 2017, class = "sf")

nyc_water <- 
  st_union(c(bx_water$geometry, 
             bk_water$geometry, 
             qn_water$geometry, 
             st_water$geometry, 
             ny_water$geometry)) %>%
  st_transform(crs = wgs84_crs)




# NYC School Zones - at https://data.cityofnewyork.us/Education/2017-2018-School-Zones/ghq4-ydq4
nyc_hs_zones <- read_sf("https://data.cityofnewyork.us/resource/9hw3-gi34.geojson") %>% 
  st_transform(., crs = wgs84_crs)

nyc_ms_zones <- read_sf("https://data.cityofnewyork.us/resource/jxpn-gg5q.geojson") %>% 
  st_transform(., crs = wgs84_crs)

nyc_es_zones <- read_sf("https://data.cityofnewyork.us/resource/xehh-f7pi.geojson") %>% 
  st_transform(., crs = wgs84_crs)

# NYC bus routes
bus_url <- "http://faculty.baruch.cuny.edu/geoportal/data/nyc_transit/may2018/bus_routes_nyc_may2018.zip"


download.file(bus_url, "data/downloads/bus_routes_nyc_may2018.zip", cacheOK = T)
unzip("data/downloads/bus_routes_nyc_may2018.zip", exdir = "data/bus_routes", overwrite = T)
nyc_bus_routes <- st_read("data/bus_routes/bus_routes_nyc_may2018.shp")
nyc_bus_routes <- st_transform(nyc_bus_routes, crs = wgs84_crs)

# NYC bus shelters
nyc_bus_shelters <- 
  st_read("https://data.cityofnewyork.us/api/geospatial/qafz-7myz?method=export&format=GeoJSON") %>% 
  st_transform(., crs=wgs84_crs)

nyc_bus_shelters$location <- str_to_title(nyc_bus_shelters$location, locale = "en")


# NYC Neighborhood Tabulation Areas (NTAs):
nyc_ntas <- st_read("https://data.cityofnewyork.us/resource/93vf-i5bz.geojson") %>% 
  st_transform(., crs = wgs84_crs)


manhattan_sf <- nyc_ntas %>% filter(boroname == "Manhattan") %>% st_union()
bronx_sf <- nyc_ntas %>% filter(boroname == "Bronx") %>% st_union()
brooklyn_sf <- nyc_ntas %>% filter(boroname == "Brooklyn") %>% st_union()
queens_sf <- nyc_ntas %>% filter(boroname == "Queens") %>% st_union()
staten_sf <- nyc_ntas %>% filter(boroname == "Staten Island") %>% st_union()

nyc_sf <- st_union(c(manhattan_sf, bronx_sf, queens_sf, brooklyn_sf, staten_sf))

# set a bbox for common views of manhattan and bronx together
man_bx_bbox <- st_bbox(st_union(manhattan_sf, bronx_sf))






# HUM II RECRUITMENT DATA ------------------------------------------------------
# This project requires the spreadsheet "historical_recruitment_data.xlsx" - HUM
# II recruitment data.

# Read in the data from each sheet
sheet_1 <- read_xlsx("data/historical_recruitment_data.xlsx", sheet = 1)
sheet_2 <- read_xlsx("data/historical_recruitment_data.xlsx", sheet = 2)
sheet_3 <- read_xlsx("data/historical_recruitment_data.xlsx", sheet = 3)

# The below three commands clean and standardize the colnames for each sheet's
# vars
str_to_lower(str_replace_all(names(sheet_1), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "") -> names(sheet_1)

str_to_lower(str_replace_all(names(sheet_2), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "") -> names(sheet_2)

str_to_lower(str_replace_all(names(sheet_3), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "") -> names(sheet_3)


# Select just the cols we need - location and application pd. Also commonly
# format zip codes (sheet two was numeric, sheets 1 and 3 were strings with
# trailing nonsignificant 0)
sheet_1 %>% mutate(studentaddress_zip =
                     str_sub(as.character(studentaddress_zip),
                             1, 5)) -> sheet_1_trimmed

sheet_2 %>% mutate(studentaddress_zip = 
                     str_sub(as.character(studentaddress_zip),
                             1, 5)) -> sheet_2_trimmed

sheet_3 %>% mutate(studentaddress_zip =
                     str_sub(as.character(studentaddress_zip),
                             1, 5)) -> sheet_3_trimmed

# Combine the three sheets
bind_rows(sheet_1_trimmed, 
          sheet_2_trimmed, 
          sheet_3_trimmed) -> recruitment_data # used to be recruitment_data_man_bx_merged

# Clean up unneeded vars
rm(list = c("sheet_1", 
            "sheet_1_trimmed", 
            "sheet_2", 
            "sheet_2_trimmed", 
            "sheet_3", 
            "sheet_3_trimmed"))


# Strip extraneous year from enrollment period and rename to "year", rename
# studentaddress_zip to zip, then cast year as numeric
recruitment_data %<>% 
  mutate(enrollment_period = str_sub(enrollment_period, 1, 4)) %>% 
  rename(year = enrollment_period, zip = studentaddress_zip) %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(!is.na(studentaddress_coordinates)) %>% # to remove one pesky record in 2018...
  separate(studentaddress_coordinates, c("lat", "long"), sep = ",") %>% 
  mutate(long = as.numeric(long), lat = as.numeric(lat))

# One hand fix to correct a typo in data entry
recruitment_data[recruitment_data$zip == "1.047", "zip"] <- "10473"


# recruitment data for manhattan and bx only
recruitment_data_man_bx <- recruitment_data %>% 
  filter(zip %in% seq(10000, 10299) | zip %in% seq(10400, 10499))

# convert recruitment_data to simple features file
recruitment_data <- recruitment_data %>% st_as_sf(coords = c("long", "lat"), crs = wgs84_crs)

# SF of application points
application_points <- recruitment_data %>% 
  mutate(registered = ifelse(!is.na(registration_completed_date), T, F))

# SF of registration points - just those that filtered as registered
registration_points <- application_points %>% 
  dplyr::select(-registration_completed_date) %>% filter(registered == T) %>% 
  dplyr::select(-registered)

```




## Special Groups Recruitment
#### Question: If I want to to target and recruit ells, where would the most sensible places be to work on targeted recruitment?

One question posed was what strategies could be employed to better target ELL recruitment efforts. 
This report reviewed the data from NYC's Department of Education on enrollment statistics in `r length(unique(districts_demo_snapshot_1718$school_dist))` districts, focusing on those in Manhattan and The Bronx. 

```{r}
# Percentile formatting for text below and to set percentile for slicing data below
schools_percentile <- 0.95

if((schools_percentile * 100) %% 10 == 1) {
  sch_pctile_ord <- paste(as.character(schools_percentile * 100), "st", sep="")
} else if ((schools_percentile * 100) %% 10 == 2) {
  sch_pctile_ord <- paste(as.character(schools_percentile * 100), "nd", sep="")
} else if ((schools_percentile * 100) %% 10 == 3) {
  sch_pctile_ord <- paste(as.character(schools_percentile * 100), "rd", sep="")
} else {
  sch_pctile_ord <- paste(as.character(schools_percentile * 100), "th", sep="")
}
```

Below is the distribution of English Language Learner recruitment among districts and schools in New York, focusing on Manhattan and the Bronx. The second plot highlights those schools recruiting ELL students at the `r paste(sch_pctile_ord)` percentile relative to all schools in New York.
```{r}

tmap_mode("plot")

plot1 <- tm_shape(nyc_area_zips, bbox = c(manhattan_sf, bronx_sf)) +
  tm_fill(col = "grey90") + 
  tm_shape(districts_demo_snapshot_1718, point.per = "feature") + 
  tm_borders() + 
  tm_fill(col = "pct_ell", title = "% ELL Enrollment, 2017-18", palette = "BuGn") +
  tm_text("school_dist") +
  tm_layout(legend.position = c("left", "top"),
            title = "Percentage of ELL enrollment by district, 2017-18",
            title.bg.color = "white", title.bg.alpha = 0.6)

# Plot schools in determined percentile for ELL enrollment in Manhattan and The Bronx:
high_ell_schools_man_bx <- schools_demo_snapshot_1718 %>% 
  filter(pct_ell > quantile(.$pct_ell, schools_percentile)) %>% 
  filter(boro %in% c("X", "M")) %>%
  .[, c("dbn", "school_name", "total_enrollment", "pct_ell")]

plot2 <- tm_shape(nyc_area_zips, bbox = c(manhattan_sf, bronx_sf)) + 
  tm_fill(col = "grey90") +
  tm_shape(nyc_sds) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(high_ell_schools_man_bx) + 
  tm_symbols(col = "red",
             size = "pct_ell",
             title.size = "% ELL enrollment",
             perceptual = T,
             sizes.legend = c(40,60,80,100), legend.size.is.portrait = F) + 
  tm_layout(legend.position = c("right", "bottom"),
            legend.bg.color = "white", legend.bg.alpha = 1, legend.frame = "black",
            title = paste(sch_pctile_ord, 
                     "percentile ELL-enrolling schools, Manhattan/Bronx"),
            title.position = c("center", "top"),
            title.bg.color = "white", title.bg.alpha = 0.6)

tmap_arrange(plot1, plot2, ncol = 2, asp = 1)
```


#### Limited English Proficiency Households
One angle for targetting recruitment efforts is examining the distribution of "limited English proficiency" households in New York. The [American Community Survey](https://www.census.gov/programs-surveys/acs/news/data-releases.html?# "American Community Survey 2017 Data Releases") tabulates limited English proficiency households by census tract, which are defined as households that do not solely speak English, in which at least one member is reported by the respondent to speak English "not very well."
```{r, lep_C16002_setup}

# GET 2017 5YR ACS HOUSEHOLD LANGUAGE SPOKEN BY LIMITED ENGLISH SPEAKING STATUS
options(tigris_use_cache = TRUE)
lep_hh_by_tract_C16002 <- 
  get_acs(geography = "tract", year = 2017, state = "NY", 
          county = c("New York", "Bronx", "Richmond", "Queens", "Kings"),
          output = "wide",
          geometry = T,
          variables = c(total_hh = "C16002_001E", 
                        spanish_lep = "C16002_004E",
                        indeur_lep = "C16002_007E",
                        asnpacisl_lep = "C16002_010E",
                        otherlang_lep = "C16002_013E",
                        total_pop = "B01003_001E")) %>% 
  st_transform(crs = wgs84_crs) %>% 
  st_erase(., nyc_water)


# break out the data from the sf object and summarize across rows for each
# language's limited english proficiency respondents. Make sure to not introduce
# NaN's by dividing by zero!
total_lep_by_tract <- lep_hh_by_tract_C16002 %>% as_tibble() %>% 
  
  # Filter out first percentile by pop tracts (often parks and cemeteries)
  filter(total_pop >= quantile(total_pop, 0.1)) %>% 
  
  # get a variable summing LEP counts of each language group as total_lep
  transmute(total_lep = select(., spanish_lep, indeur_lep, 
                               asnpacisl_lep, otherlang_lep) %>% rowSums(),
            
            # and one for percentage. If total_lep is 0, hard enter 0% for pct
            pct_lep = ifelse(total_lep == 0, 0, 100 * total_lep / total_hh), 
            
            #keep GEOID variable
            GEOID = GEOID)

# merge the data back into the geometry
lep_hh_by_tract_C16002 <- inner_join(lep_hh_by_tract_C16002, 
                                    total_lep_by_tract, 
                                    by = c("GEOID" = "GEOID"))

# prep for plotting: add school districts for each tract.
lep_hh_by_tract_C16002 <- st_join(lep_hh_by_tract_C16002, nyc_sds)


```
The below interactive map shows the percentage of limited English proficiency households per census tract in NYC, with school district boundaries overlaid. __Note that this calculation has taken out of consideration those tracts with extremely low (first percentile and below) total tract populations.__ While some of those tracts have high proportions of limited English proficiency households, the total number of people living in them is extremely small, and most of them represent land that is largely a park or cemetery.

```{r, LEP_household_proportions_map}

# Plot map of LEP household proportions by tract
tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
tm_shape(nyc_sds, point.per = "feature", bbox = man_bx_bbox) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
tm_shape(lep_hh_by_tract_C16002) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep"))

# # Hidden: Plot mode available for printed version of report.
# tmap_mode("plot")
# tm_shape(nyc_area_zips, bbox = c(manhattan_sf, bronx_sf)) + tm_fill(col = "grey90") +
# tm_shape(lep_tracts) +
#   tm_fill("pct_lep", palette = "Reds", style = "pretty",
#           title = "% LEP households\nby census Tract") +
#   tm_borders(alpha = 0.1) +
#   tm_credits("Source: 2013-2017 ACS Five-Year Estimates",
#              bg.color = "white", bg.alpha = 0.7, position = c("left", "bottom")) +
#   tm_layout(legend.position = c("left", "top"),
#             legend.height = .4,
#             legend.width = .8,
#             legend.title.size = 1,
#             legend.title.fontface = "bold
```



```{r }


pct <- 0.75
lep_hh_manbx_by_quantile <- 
  lep_hh_by_tract_C16002 %>% filter(pct_lep >= quantile(.$pct_lep, pct)) %>% 

  # need to add a way to filter boros

# Get all bus shelters in NYC that are in pctile-th-percentile tracts by proportion
# of LEP households
bus_shelters_in_high_lep_tracts <- 
  nyc_bus_shelters[lep_hh_quantile_subset, ]

bus_shelters_in_high_lep_tracts_manbx <- bus_shelters_in_high_lep_tracts %>% 
  filter(boro_name %in% c("Bronx", "Manhattan"))

tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
tm_shape(nyc_sds, point.per = "feature", bbox = man_bx_bbox) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
tm_shape(lep_hh_quantile_subset) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep")) +
tm_shape(bus_shelters_in_high_lep_tracts) +
  tm_dots(col = "#42f4e2")


```
The following `r as.character(nrow(bus_shelters_in_high_lep_tracts))` bus shelters are located in the tracts in Manhattan and the Bronx with the highest (`r as.character(pct * 100)`th percentile) proportions of limited english proficiency. Again, census tracts with extremely low population (often parks, cemeteries and other such areas) have been removed prior to calculation.

`r high_lep_bus_shelters_manbx`

For targeting these areas, two options stand out: mailers to dense housing, and bus shelter/bus line advertisements in these areas.

