---
title: "High School Recruitment Report"
author: "Ryder Cobean"
output: 
  output: html_document
---

<style>
.leaflet {
    margin: auto;
}
</style>



This report attempts to answer a number of questions for a School in the South Bronx, New York City. These analyses were done based off of data from that school. All data had no individual student identifiers. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', fig.align="center", echo = F, warning = F, message = F)

library(sf)
library(leaflet)
library(revgeo)
library(rmarkdown)
library(kableExtra)
library(magrittr)
library(viridis)
library(viridisLite)
library(tmaptools)
library(tmap)
library(ggplot2)
library(readxl)
library(tigris)
library(tidycensus)
library(acs)
library(tidyverse)
library(BAMMtools) # for getting jenks natural breaks
```


```{r data_prep, results='hide', cache=T}

if(!dir.exists("data")) {
  dir.create("data/downloads", recursive = T)
} else if (!dir.exists("data/downloads")) {
  dir.create("data/downloads")
}


# if(Sys.getenv("CENSUS_API_KEY") == "") {
#   
#   census_api_key(getPass("Enter U.S. census API key to install: "), install = T, overwrite = T)
#   readRenviron("~/.Renviron")
#   
# }


##### SET A REFERENCE VARIABLE HERE: CRS FOR ALL LAYERED PLOTS ##### use this
# CRS to transform projeections of any other layer being used when necessary

wgs84_crs <-  "+proj=longlat +datum=WGS84 +no_defs"



# Download shapefiles for NYC area ZCTAs to render as background layer
options(tigris_use_cache = T)
nyc_area_zips <- 
  zctas(cb = T, starts_with = c('070','071','072','073','074',
                                '075','076','100','101','102',
                                '103','104','105','106','107',
                                '108','109','110','111','112',
                                '113','114','115','116')) %>% 
  as("sf") %>%
  st_transform(crs = wgs84_crs)


# get a simple block of NJ coastline to add to baselayers of maps
nj_land <- nyc_area_zips %>% filter(str_detect(GEOID10, "^07")) %>% st_union
  

# Filter just the ZCTAs we're interested in - Man/Bx
man_bx_zips <- nyc_area_zips %>% filter(str_detect(GEOID10, "^100|^101|^102|^103|^104"))


# Get NYCHA location geojson
nycha_locations <- 
  st_read("https://data.cityofnewyork.us/resource/5j2e-zhmb.geojson", stringsAsFactors = F) %>% 
  st_transform(crs = wgs84_crs)


# NYC's SD geojson file only pulls in data as characters, which leads sf to read
# them as factors which breaks all attempts at analysis. forcing NOT to factors
# brings everything in as characters, which lets us cast each attribute to
# numeric
nyc_sds <- st_read("https://data.cityofnewyork.us/resource/cuae-wd7h.geojson", stringsAsFactors = F) %>% 
  st_transform(crs = wgs84_crs) %>% 
  dplyr::select(-shape_area, -shape_leng) %>% 
  mutate(school_dist = as.integer(school_dist)) %>% 
  # mutate(school_dist = as.factor(school_dist)) %>% 
  group_by(school_dist) %>% 
  summarize()




districts_demo_snapshot <- 
  read_csv("https://data.cityofnewyork.us/resource/dndd-j759.csv") %>% 
  rename_all(
    funs(
      str_to_lower(.) %>% 
      str_replace_all(., "\\s", "_") %>% 
      str_replace_all(., "#", "num") %>% 
      str_replace_all(., "%", "pct") %>% 
      str_replace_all(., "\\(|\\)", "") %>% 
      str_replace_all(., "\\&", "") %>% 
      str_replace_all(., "_{2,}", "_")
    )
  ) %>% 
  mutate(administrative_district = as.integer(administrative_district)) %>% 
  # mutate(administrative_district = as.factor(administrative_district)) %>% 
  rename(num_swd = students_with_disabilities_1, 
         pct_swd = students_with_disabilities_2,
         num_ell = english_language_learners_1, 
         pct_ell = english_language_learners_2, 
         num_pov = poverty_1, 
         pct_pov = poverty_2) %>% 
  inner_join(nyc_sds, ., 
             by = c("school_dist" = "administrative_district"))

districts_demo_snapshot_1718 <- districts_demo_snapshot %>% 
  filter(year == "2017-18")


# Shapefile of public school points - This is an ESRI shape file of school point
# locations based on the official address.  It includes some additional basic
# and pertinent information needed to link to other data sources. It also
# includes some basic school information such as Name, Address, Principal, and
# Principal’s contact information.

download.file(url = "https://data.cityofnewyork.us/download/jfju-ynrr/application%2Fzip", destfile = "data/downloads/school_points.zip")
unzip("data/downloads/school_points.zip", exdir = "data/school_points", overwrite = T)

school_points <- st_read("data/school_points/Public_Schools_Points_2011-2012A.shp") %>% 
  rename_all(str_to_lower) %>% 
  # clean up trailing unicode characters in ats_code
  mutate(ats_code = str_sub(ats_code, 1, 6)) %>% 
  mutate(principal = str_to_title(principal, locale = "en")) %>% 
  st_transform(school_points, crs = wgs84_crs)

# School demographic data from NYC gov't. Will merge with above points data.
schools_demo_snapshot <- 
  read_csv("https://data.cityofnewyork.us/api/views/s52a-8aq6/rows.csv?accessType=DOWNLOAD&bom=true&format=true") %>% 
  rename_all(
    funs(
      str_to_lower(.) %>% 
      str_replace_all(., "\\s", "_") %>% 
      str_replace_all(., "#", "num") %>% 
      str_replace_all(., "%", "pct") %>% 
      str_replace_all(., "\\(|\\)", "") %>% 
      str_replace_all(., "\\&", "") %>% 
      str_replace_all(., "_{2,}", "_")
    )
  ) %>% 
  rename(num_swd = num_students_with_disabilities, 
         pct_swd = pct_students_with_disabilities,
         num_ell = num_english_language_learners, 
         pct_ell = pct_english_language_learners,
         num_pov = num_poverty, 
         pct_pov = pct_poverty) %>% 
  inner_join(school_points, ., by = c('ats_code' = 'dbn'))
  
# filter out old years of schools demo snapshot
schools_demo_snapshot_1718 <- schools_demo_snapshot %>% filter(year == "2017-18")




# The below sets up the necessary shapefile and function to clip any shapefile
# to the NYC shoreline (for example, the US Census Tract boundaries, which
# extend into the Hudson and East Rivers). Passing your shapefile and nyc_water
# to st_erase will clip them to the shoreline. Curiously, 
st_erase <- function(x, y) {
  st_difference(x, st_union(st_combine(y)))
}

bx_water <- tigris::area_water("NY", county = "Bronx", class = "sf")
bk_water <- tigris::area_water("NY", county = "Kings", class = "sf")
qn_water <- tigris::area_water("NY", county = "Queens", class = "sf")
st_water <- tigris::area_water("NY", county = "Richmond", class = "sf")
ny_water <- tigris::area_water("NY", county = "New York", year = 2017, class = "sf")

nyc_water <- 
  st_union(c(bx_water$geometry, 
             bk_water$geometry, 
             qn_water$geometry, 
             st_water$geometry, 
             ny_water$geometry)) %>%
  st_transform(crs = wgs84_crs)


# Get tracts (unclipped to shoreline)
trct <- 
  tigris::tracts(state = "NY", 
                 county = c('Bronx',
                            'Kings',
                            'New York',
                            'Queens',
                            'Richmond')) %>% 
  as("sf") %>%  # cast to sf
  st_transform(crs=wgs84_crs) %>% 
  
  # make easier to work with with county name
  mutate(countyname = recode(.$COUNTYFP, 
                             `005`="Bronx", 
                             `047`="Kings", 
                             `061`="New York", 
                             `081`="Queens", 
                             `085`="Richmond")) %>% 
  
  mutate(tractname = paste(NAMELSAD, ", ",
                           countyname,
                           " County",
                           sep = ""))

nyc_tracts <- st_erase(trct, nyc_water) # erase water from edges 


# NYC School Zones - at https://data.cityofnewyork.us/Education/2017-2018-School-Zones/ghq4-ydq4
nyc_hs_zones <- read_sf("https://data.cityofnewyork.us/resource/9hw3-gi34.geojson") %>% 
  st_transform(., crs = wgs84_crs)

nyc_ms_zones <- read_sf("https://data.cityofnewyork.us/resource/jxpn-gg5q.geojson") %>% 
  st_transform(., crs = wgs84_crs)

nyc_es_zones <- read_sf("https://data.cityofnewyork.us/resource/xehh-f7pi.geojson") %>% 
  st_transform(., crs = wgs84_crs)

# NYC bus routes
bus_url <- "http://faculty.baruch.cuny.edu/geoportal/data/nyc_transit/may2018/bus_routes_nyc_may2018.zip"


download.file(bus_url, "data/downloads/bus_routes_nyc_may2018.zip", cacheOK = T)
unzip("data/downloads/bus_routes_nyc_may2018.zip", exdir = "data/bus_routes", overwrite = T)
nyc_bus_routes <- st_read("data/bus_routes/bus_routes_nyc_may2018.shp")
nyc_bus_routes <- st_transform(nyc_bus_routes, crs = wgs84_crs)


# merge bus routes by route (directions currently their own shapes)
nyc_bus_routes %<>% group_by(route_id, route_shor, route_long, color) %>% 
  summarize() %>% ungroup()

# NYC bus shelters
nyc_bus_shelters <- 
  st_read("https://data.cityofnewyork.us/api/geospatial/qafz-7myz?method=export&format=GeoJSON") %>% 
  st_transform(., crs=wgs84_crs)

# nyc_bus_shelters$location <- str_to_title(nyc_bus_shelters$location, locale = "en")

nyc_bus_shelters %<>% mutate(location = str_to_title(nyc_bus_shelters$location, locale = "en"),
                            coords = paste(latitude, longitude, sep = ", "))


# NYC Neighborhood Tabulation Areas (NTAs):
nyc_ntas <- st_read("https://data.cityofnewyork.us/resource/93vf-i5bz.geojson") %>% 
  st_transform(., crs = wgs84_crs)


manhattan_sf <- nyc_ntas %>% filter(boroname == "Manhattan") %>% st_union()
bronx_sf <- nyc_ntas %>% filter(boroname == "Bronx") %>% st_union()
brooklyn_sf <- nyc_ntas %>% filter(boroname == "Brooklyn") %>% st_union()
queens_sf <- nyc_ntas %>% filter(boroname == "Queens") %>% st_union()
staten_sf <- nyc_ntas %>% filter(boroname == "Staten Island") %>% st_union()

nyc_sf <- st_union(c(manhattan_sf, bronx_sf, queens_sf, brooklyn_sf, staten_sf))

nyc_area_land <- st_union(nj_land, nyc_sf)

# set a bbox for common views of manhattan and bronx together
man_bx_bbox <- st_bbox(st_union(manhattan_sf, bronx_sf))






# HUM II RECRUITMENT DATA ------------------------------------------------------
# This project requires the spreadsheet "historical_recruitment_data.xlsx" - HUM
# II recruitment data.

# Read in the data from each sheet
sheet_19 <- read_xlsx("schooldata/historical_recruitment_data_0319.xlsx")
sheet_18 <- read_xlsx("schooldata/historical_recruitment_data.xlsx", sheet = 2)
sheet_17 <- read_xlsx("schooldata/historical_recruitment_data.xlsx", sheet = 3)

# The below three commands clean and standardize the colnames for each sheet's
# vars

names(sheet_17) <- 
  str_to_lower(str_replace_all(names(sheet_17), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "")

names(sheet_18) <- 
  str_to_lower(str_replace_all(names(sheet_18), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "")

names(sheet_19) <- 
  str_to_lower(str_replace_all(names(sheet_19), 
                             "[^[:alnum:]]|[^[:ascii:]]", "_")) %>% 
  str_replace_all("[:punct:]+", "_") %>% 
  str_replace_all("^[:punct:]|[:punct:]$", "")


# Select just the cols we need - location and application pd. Also commonly
# format zip codes (sheet two was numeric, sheets 1 and 3 were strings with
# trailing nonsignificant 0)
sheet_19_trimmed <- 
  sheet_19 %>% mutate(studentaddress_zip =
                     str_sub(as.character(studentaddress_zip), 1, 5))
sheet_18_trimmed <- 
  sheet_18 %>% mutate(studentaddress_zip = 
                     str_sub(as.character(studentaddress_zip), 1, 5))

sheet_17_trimmed <- 
  sheet_17 %>% mutate(studentaddress_zip =
                     str_sub(as.character(studentaddress_zip), 1, 5))

# Combine the three sheets

recruitment_data <- 
  bind_rows(sheet_17_trimmed, 
            sheet_18_trimmed, 
            sheet_19_trimmed) # used to be recruitment_data_man_bx_merged

# Clean up unneeded vars
rm(list = c("sheet_1", 
            "sheet_1_trimmed", 
            "sheet_2", 
            "sheet_2_trimmed", 
            "sheet_3", 
            "sheet_3_trimmed"))


# Strip extraneous year from enrollment period and rename to "year", rename
# studentaddress_zip to zip, then cast year as numeric
recruitment_data %<>% 
  mutate(enrollment_period = str_sub(enrollment_period, 1, 4)) %>% 
  rename(year = enrollment_period) %>% 
  filter(!is.na(studentaddress_coordinates)) %>% # to remove one pesky record in 2018...
  separate(studentaddress_coordinates, c("lat", "long"), sep = ",") %>% 
  mutate(long = as.numeric(long), lat = as.numeric(lat)) %>% 
  
  # convert to sf and intersect with tracts, zips, and ntas shapefiles
  st_as_sf(coords = c("long", "lat"), 
           crs    = wgs84_crs) %>% 
  st_intersection(., nyc_area_zips) %>% 
  st_intersection(., nyc_ntas) %>% 
  st_intersection(., nyc_tracts) %>% 
  st_intersection(., nyc_sds) %>% 
  
  # rename and reclass key geography variables
  mutate(year        = as.integer(year),
         district    = as.integer(district),
         zip         = as.integer(GEOID10),
         tractname   = as.factor(paste(NAMELSAD, boroname, sep = ", "))) %>% 
  
  # add a logical variable for whether applicant was registered or not
  mutate(registered = ifelse(!is.na(registration_completed_date), T, F))


# One hand fix to correct a typo in data entry
# recruitment_data[recruitment_data$zip == "1.047", "zip"] <- "10473"

# recruitment data for manhattan and bx only
recruitment_data_man_bx <- recruitment_data %>% 
  filter(boroname %in% c("Bronx", "Manhattan"))
  # filter(zip %in% seq(10000, 10299) | zip %in% seq(10400, 10499))

# SF of application points
application_points <- recruitment_data # %>% 
  # mutate(registered = ifelse(!is.na(registration_completed_date), T, F))

# SF of registration points - just those that filtered as registered
registration_points <- application_points %>% 
  dplyr::select(-registration_completed_date) %>% filter(registered == T) %>% 
  dplyr::select(-registered)

```


```{r functions, results='hide', cache=T}

# Used in bus route chunk.

st_intersects_tally_attr <-
  function(x = NULL,
           y = NULL,
           attribute = NULL) {
    require(sf)
    
    # a <- attribute
    a <- attribute
    
    
    # a list of x's length with each record containing a vector of indices of y's
    # records that intersect with that record in x
    xy <- st_intersects(x, y)
    
    
    attr_n <- c() # empty vector to fill
    for (r in 1:length(xy)) {
      n <- 0
      
      if (length(xy[[r]]) == 0) {
        n <- 0
        
      } else {
        for (i in 1:length(xy[[r]])) {
          n <- n + st_drop_geometry(y)[i, a]
          
        }
        
      }
      
      attr_n <- c(attr_n, n)
      
    }
    
    
    return(x %>% bind_cols((!!a) := attr_n))
    
  }
```


## Historical Recruitment Data

```{r historical_recruitment}

# create a tidy set of application totals of this with all years
recruitment_by_tract_tidy <- 
  recruitment_data %>% 
  st_drop_geometry() %>% 
  count(GEOID, year, registered) %>% 
  complete(GEOID, year, registered, fill = list(n = 0)) %>% 
  mutate(pct_total = n / sum(n) * 100) %>% 
  group_by(year) %>% 
  mutate(pct_yr = n / sum(n) * 100) %>% 
  ungroup() %>% 
  inner_join(nyc_tracts, ., by = c("GEOID" = "GEOID"))


recruitment_by_sd_tidy <- 
  recruitment_data %>% 
  st_drop_geometry %>% 
  count(school_dist, year, registered) %>% 
  complete(school_dist, year, registered, fill = list(n = 0)) %>% 
  mutate(pct_total = n / sum(n)) %>% 
  group_by(year) %>% 
  mutate(pct_yr = n / sum(n) * 100) %>% 
  ungroup() %>% 
  inner_join(nyc_sds, ., by = c("school_dist" = "school_dist"))

applications_by_tract_yearly_totals <- 
  recruitment_by_tract_tidy %>% 
  group_by(GEOID, tractname, year) %>% 
  summarize(n = sum(n)) %>% 
  group_by(year) %>% 
  mutate(pct_yr = n / sum(n) * 100) %>% 
  ungroup() %>% 
  mutate(pct_total = n / sum(n) * 100) %>%
  filter(n > 0)

apps_jenks_breaks <- applications_by_tract_yearly_totals$n %>% getJenksBreaks(6)

registrations_by_tract_yearly_totals <-
  recruitment_by_tract_tidy %>%
  filter(registered == T, n > 0)

regs_jenks_breaks <- getJenksBreaks(registrations_by_tract_yearly_totals$n, 6)

```

### Application Data {.tabset}

Below are the distributions of applications received from the `r applications_by_tract_yearly_totals$year %>% min()`-`r (applications_by_tract_yearly_totals$year %>% min() + 1)` to the `r applications_by_tract_yearly_totals$year %>% max()`-`r applications_by_tract_yearly_totals$year %>% max() + 1` school years, arranged by NYC Census Tract. In total, `r applications_by_tract_yearly_totals$n %>% sum()` applications were received from `r applications_by_tract_yearly_totals$tractname %>% unique() %>% length()` tracts.


```{r applications_plot}

tmap_mode("view")

a_2017 <- 
  tm_tiles(providers$CartoDB.PositronNoLabels) + 
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(applications_by_tract_yearly_totals %>% 
               filter(year == 2017)) + 
      tm_borders(alpha = 0.2) +
      tm_fill(col = "n",
            id="tractname",
            breaks = apps_jenks_breaks,
            title = "2017 Apps",
            popup.vars = c("Applications"="n",
                           "% of Year's Applications"="pct_yr")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
            set.zoom.limits = c(11,14)) + 
    tm_layout(outer.margins = 20)    

a_2018 <- 
  tm_tiles(providers$CartoDB.PositronNoLabels) + 
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(applications_by_tract_yearly_totals %>% 
               filter(year == 2018)) + 
      tm_borders(alpha = 0.2) +
      tm_fill(col = "n",
            id="tractname",
            breaks = apps_jenks_breaks,
            title = "2018 Apps",
            popup.vars = c("Applications"="n",
                           "% of Year's Applications"="pct_yr")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
            set.zoom.limits = c(11,14)) + 
    tm_layout(outer.margins = 20)

a_2019 <- 
  tm_tiles(providers$CartoDB.PositronNoLabels) + 
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(applications_by_tract_yearly_totals %>% 
               filter(year == 2019)) + 
      tm_borders(alpha = 0.2) +
      tm_fill(col = "n",
            id="tractname",
            breaks = apps_jenks_breaks,
            title = "2019 Apps",
            popup.vars = c("Applications"="n",
                           "% of Year's Applications"="pct_yr")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
            set.zoom.limits = c(11,14)) + 
    tm_layout(outer.margins = 20)

```
  
#### 2017 Applications  
  
```{r}
a_2017
```

#### 2018 Applications  
  
```{r}
a_2018
```

#### 2019 Applications  
  
```{r}
a_2019
```

### Registration Data {.tabset}

Below are the distributions of applications that resulted in the student being accepted and registered. The registration data currently tracks the `r registrations_by_tract_yearly_totals$year %>% min()`-`r registrations_by_tract_yearly_totals$year %>% min() + 1` to the `r registrations_by_tract_yearly_totals$year %>% max()`-`r registrations_by_tract_yearly_totals$year %>% max() + 1` school years, with `r registrations_by_tract_yearly_totals$n %>% sum()` registrations in total occurring from `r registrations_by_tract_yearly_totals$tractname %>% unique() %>% length()` tracts in New York City.

```{r registrations_plot}

tmap_mode("view")


regs_2017 <- tm_tiles(providers$CartoDB.PositronNoLabels) +
  tm_shape(nyc_sds, point.per = "feature") +
    tm_borders(alpha = 0.5) +
    tm_text(text = "school_dist") +
  tm_shape(registrations_by_tract_yearly_totals %>%
             filter(year == 2017)) +
    tm_borders(alpha = 0.2) +

    tm_fill(col = "n",
            palette = "YlOrBr",
            id="tractname",
            breaks = regs_jenks_breaks,
            title = "2017 Regs",
            popup.vars = c("Registrations"="n",
                           "% of Year's Registrations"="pct_yr")) +
    tm_view(set.view = c(-73.86652, 40.85263, 11),
          set.zoom.limits = c(11,14))

regs_2018 <- tm_tiles(providers$CartoDB.PositronNoLabels) +
  tm_shape(nyc_sds, point.per = "feature") +
    tm_borders(alpha = 0.5) +
    tm_text(text = "school_dist") +
  tm_shape(registrations_by_tract_yearly_totals %>%
             filter(year == 2018)) +
    tm_borders(alpha = 0.2) +

    tm_fill(col = "n",
            palette = "BuGn",
            id="tractname",
            breaks = regs_jenks_breaks,
            title = "2018 Regs",
            popup.vars = c("Registrations"="n",
                           "% of Year's Registrations"="pct_yr")) +
    tm_view(set.view = c(-73.86652, 40.85263, 11),
          set.zoom.limits = c(11,14))



```
  
#### 2017 Registrations
```{r}
regs_2017
```

#### 2018 Registrations
```{r}
regs_2018
```


  
### Application to Registration Performance {.tabset}

```{r recruitment_performance_setup, include=F}

tracts_by_recruitment_performance <- 
  recruitment_data %>% 
  count(year, GEOID, registered) %>% 
  spread(key = registered, value = n, fill = 0) %>% 
  complete(year, GEOID, fill = list(`TRUE` = 0, `FALSE` = 0)) %>% 
  transmute(Year = year,
            GEOID = GEOID,
            Applied = `TRUE` + `FALSE`,
            `Not Registered` = `FALSE`,
            Registered = `TRUE`,
            `Recruitment Rate` = ifelse(Applied == 0, 0, Registered / Applied))


tracts_by_recruitment_performance <- 
  recruitment_data %>% 
  count(year, GEOID, registered) %>% 
  spread(key = registered, value = n, fill = 0) %>% 
  transmute(Year = year,
            GEOID = GEOID,
            Applied = `TRUE` + `FALSE`,
            `Not Registered` = `FALSE`,
            Registered = `TRUE`,
            `Recruitment Rate` = ifelse(Applied == 0, 0, Registered / Applied)) %>% 
  st_drop_geometry() %>% 
  inner_join(., nyc_tracts, by = c("GEOID" = "GEOID")) %>% 
  st_as_sf(.)

# Applications to Registrations Rate Map:
apps_to_regs_map <- function(years = seq(0,10000), 
                             pal = "YlOrBr") {
  
  data <- tracts_by_recruitment_performance
  years <- years
  pal <- pal

  data %<>%
    filter(Year %in% years)

  map_plot <-
    tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) +
    tm_shape(nyc_sds, point.per = "feature") +
      tm_borders(alpha = 0.7) +
      tm_text(text = "school_dist") +
    tm_shape(data) +
      tm_borders(alpha = 0.2) +
      tm_fill(col = "Recruitment Rate",
              palette = pal,
              id = "tractname",
              title =

                # logic to handle whether years renders as
                # YYYY-YY or just YYYY depending on whether
                # multiple years are called in args
                if (length(years) > 1) {
                  paste("Regs:Apps, ", "\n", min(years), "-",
                            substr(toString(max(years)), 3, 4),
                            sep = "")
                } else {
                  paste("Apps:Regs, ", "\n",
                        years,
                        sep = "")
                },

              popup.vars =
                c("Tract ID #"="GEOID",
                  "Applied",
                  "Registered",
                  "Not Registered",
                  "Recruitment Rate")) +
      tm_view(set.view = c(-73.86652, 40.85263, 11),
              set.zoom.limits = c(11,14)) +
      tm_layout(outer.margins = 20)
  
  return(map_plot)
  
}

```


Below are a number of views of the recruitment performance, by different geographies.

#### By School District, All Years

Below is the comparative performance of recruitment in each of New York's `r nyc_sds %>% length()` school district boundaries, showing the number of total applications received and the number of registrations completed from those applications, per year, for the `r min(recruitment_by_sd_tidy$year)`-`r min(recruitment_by_sd_tidy$year) + 1` to the `r max(recruitment_by_sd_tidy$year)`-`r max(recruitment_by_sd_tidy$year) + 1` school years.


```{r app_reg_performance}


ggplot(recruitment_by_sd_tidy %>% 
         mutate(school_dist = as.factor(school_dist))) + 
  geom_bar(aes(x = school_dist, y = n, fill = registered), stat = "identity") +
  ggtitle("Recruitment Performance by Applicant's School District", 
          subtitle = "School Years Beginning 2017 through 2019") +
  theme(plot.title      = element_text(hjust = 0.5),
        plot.subtitle   = element_text(hjust = 0.5),
        legend.position = "bottom") +
  xlab("School District") + 
  ylab("Count") +
  facet_grid(year ~ .)




```


#### By Census Tracts, 2017

Below are census tracts depicting rate of registration (total registrations / total applications, by applications originating in the census tract.)
```{r}
apps_to_regs_map(2017)
```

#### By Census Tracts, 2018

Below are census tracts depicting rate of registration (total registrations / total applications, by applications originating in the census tract.)
```{r}
apps_to_regs_map(2018, "BuGn")
```


### Limited English Proficiency Households

One angle for targetting recruitment efforts is examining the distribution of "limited English proficiency" households in New York. The [American Community Survey](https://www.census.gov/programs-surveys/acs/news/data-releases.html?# "American Community Survey 2017 Data Releases") tabulates limited English proficiency households by census tract, which are defined as households that do not solely speak English, in which at least one member is reported by the respondent to speak English "not very well."
```{r lep_C16002_setup}

# GET 2017 5YR ACS HOUSEHOLD LANGUAGE SPOKEN BY LIMITED ENGLISH SPEAKING STATUS
options(tigris_use_cache = TRUE)
lep_hh_by_tract_C16002 <- 
  get_acs(geography = "tract", year = 2017, state = "NY", 
          county = c("New York", "Bronx", "Richmond", "Queens", "Kings"),
          output = "wide",
          geometry = T,
          variables = c(total_hh = "C16002_001E", 
                        spanish_lep = "C16002_004E",
                        indeur_lep = "C16002_007E",
                        asnpacisl_lep = "C16002_010E",
                        otherlang_lep = "C16002_013E",
                        total_pop = "B01003_001E")) %>% 
  st_transform(crs = wgs84_crs) %>% 
  st_erase(., nyc_water)

# get County name from longform tract name in NAME, get boro name from county names
# Don't need this anymore since adding boro
lep_hh_by_tract_C16002 %<>%
  mutate(county_name = str_extract(NAME, "((?<=\\d, )[:graph:]+).+(?=\\sCounty)")) 

# break out the data from the sf object and summarize across rows for each
# language's limited english proficiency respondents. Make sure to not introduce
# NaN's by dividing by zero!
total_lep_by_tract <- lep_hh_by_tract_C16002 %>% as_tibble() %>% 
  
  # Filter out first percentile by pop tracts (often parks and cemeteries)
  filter(total_pop >= quantile(total_pop, 0.1)) %>% 
  
  # get a variable summing LEP counts of each language group as total_lep
  transmute(total_lep = dplyr::select(., spanish_lep, indeur_lep, 
                                      asnpacisl_lep, otherlang_lep) %>% rowSums(),
            
            # and one for percentage. If total_lep is 0, hard enter 0% for pct
            pct_lep = ifelse(total_lep == 0, 0, 100 * total_lep / total_hh), 
            
            #keep GEOID variable
            GEOID = GEOID)

# merge the data back into the geometry
lep_hh_by_tract_C16002 <- inner_join(lep_hh_by_tract_C16002, 
                                     total_lep_by_tract, 
                                     by = c("GEOID" = "GEOID"))

# prep for plotting: add school districts for each tract.
lep_hh_by_tract_C16002 <- st_join(lep_hh_by_tract_C16002, nyc_sds)

# filter to just Manhattan and Bronx
lep_hh_by_tract_C16002_manbx <- filter(lep_hh_by_tract_C16002,
                                       county_name %in% c("New York", "Bronx"))


```

The below interactive map shows the percentage of limited English proficiency households per census tract in NYC, with school district boundaries overlaid. __Note that this calculation has taken out of consideration those tracts with extremely low (first percentile and below) total tract populations.__ While some of those tracts have high proportions of limited English proficiency households, the total number of people living in them is extremely small, and most of them represent land that is largely a park or cemetery.

```{r LEP_household_proportions_map}

# Plot map of LEP household proportions by tract
tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  tm_shape(nyc_sds, point.per = "feature", bbox = man_bx_bbox) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  tm_shape(lep_hh_by_tract_C16002) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep"))



```



```{r bus_shelters_lep_setup, cache=T, dependson=c("setup", "functions"), include=FALSE}
pct <- 0.80
high_lep_hh_manbx <- 
  lep_hh_by_tract_C16002_manbx %>% 
  filter(pct_lep >= quantile(.$pct_lep, pct))


# Get all bus shelters in NYC that are in pctile-th-percentile tracts by proportion
# of LEP households


# 
# high_lep_bus_shelters_manbx <- 
#   nyc_bus_shelters[high_lep_hh_manbx, ]

high_lep_bus_shelters_manbx <- st_intersection(nyc_bus_shelters, high_lep_hh_manbx)

# Prep a version of the dataframe of 
shelter_table <- high_lep_bus_shelters_manbx %>% as_tibble() %>% 
  transmute(`Shelter ID` = shelter_id, Coordinates = coords,
            Location = paste(location, " at ", str_to_title(at_between), sep=""),
            `Census Tract` = GEOID,
            `LEP Households` = total_lep) %>% 
  arrange(desc(`LEP Households`))

```

### Bus Shelters in High-LEP Census Tracts
The following `r as.character(nrow(high_lep_bus_shelters_manbx))` bus shelters are located in the tracts in Manhattan and the Bronx with the highest (`r as.character(pct * 100)`th percentile) proportions of limited english proficiency. Again, census tracts with extremely low population (often parks, cemeteries and other such areas) have been removed prior to calculation.


```{r bus_shelters_lep_plot}
tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  tm_shape(nyc_sds, point.per = "feature", 
           bbox = st_bbox(filter(high_lep_bus_shelters_manbx, boro_name == "Bronx"))) + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  tm_shape(high_lep_hh_manbx) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          title = "% LEP Households",
          id = "NAME",
          showNA = F,
          popup.vars = c("Tract ID"="GEOID", "Name"="NAME", 
                         "School District"="school_dist",
                         "Total Households"="total_hh",
                         "% LEP Households"="pct_lep")) +
  tm_shape(high_lep_bus_shelters_manbx) +
  tm_dots(col = "#42f4e2",
          id = "shelter_id",
          popup.vars = c("Location" = "location",
                         "Coordinates" = "coords",
                         "Nearby LEP Households" = "total_lep",
                         "% LEP households in Tract" = "pct_lep"))

```

\  
\  


```{r}
kable(shelter_table, caption = "Bus shelters in proportionally high-LEP household tracts") %>%
  kable_styling(bootstrap_options = "striped") %>% 
  scroll_box(height = "300px")

```
\  

The entire table can be copied and pasted into a spreadsheet and will retain its formatting.
\  

```{r bus_routes}


# Create a subset of buses that just run in Manhattan and the Bronx
manbx_bus_sgbp <- st_intersects(nyc_bus_routes, c(manhattan_sf, bronx_sf))
manbx_bus_routes <- nyc_bus_routes[lengths(manbx_bus_sgbp) > 0, ]
rm(manbx_bus_sgbp)  # get rid of helper variable

# Tally the value "total_lep" for its intersects with bus routes passing through
# tracts with the attribute
manbx_bus_routes_with_lep <- st_intersects_tally_attr(manbx_bus_routes,
                                                      lep_hh_by_tract_C16002_manbx,
                                                      "total_lep")



manbx_bus_routes_with_lep %<>% 
  top_n(10, total_lep) %>% 
  arrange(desc(total_lep)) %>% 
  mutate(route_id = fct_drop(route_id))


```

### Bus Routes Along High-LEP Areas

In addition to advertising on the bus stops located within the highest-proportion LEP census tracts in New York, placing advertisements for the school on the bus lines passing near the most LEP households is another strategy. Below are the top `r nrow(manbx_bus_routes_with_lep)` routes in Manhattan and the Bronx, by the number of Low English Proficiency households they pass by (total LEP household counts of the census tracts they pass through):
  
  
  ```{r bus_route_lep_plot}


tmap_mode("view")
tm_basemap(leaflet::providers$CartoDB.PositronNoLabels) + 
  
  #  NYC School Districts Base - SD Borders
  tm_shape(nyc_sds, point.per = "feature") + 
  tm_borders(alpha = 0.8) + 
  tm_text("school_dist") +
  
  #  Polygons of high proportion Low English Proficiency Tracts  
  tm_shape(high_lep_hh_manbx) + 
  tm_borders(alpha = 0.2) + 
  tm_fill(col = "pct_lep", 
          palette = "Reds",
          # title = "% LEP Households",
          id = "NAME",
          legend.show = F,
          showNA = F,
          alpha = 0.5,
          popup.vars = c("Tract ID: "="GEOID", 
                         "Name: "="NAME", 
                         "School District: "="school_dist",
                         "Total Households: "="total_hh",
                         "% LEP Households: "="pct_lep")) +
  
  #  Top n bus routes that pass through the census tracts with the most LEP
  #  households
  tm_shape(manbx_bus_routes_with_lep) + 
  tm_lines(lwd = 3,
           id = "route_id",
           col = "route_id",
           palette = "Accent",
           title.col = "Route",
           # col= "#2eaa9e",
           alpha = 1,
           popup.vars = c("Route ID: " = "route_id",
                          "Name: " = "route_long",
                          "# LEP Households: " = "total_lep")) +
  
  #  Focus view to the center of the Bronx, set zoom to fill view with Bronx
  tm_view(set.view=c(-73.88251, 40.83489, 12),
          set.zoom.limits = c(11, 15))



```
\  
\  
```{r bus_route_lep_table}

manbx_bus_routes_with_lep_kable <- manbx_bus_routes_with_lep %>% 
  dplyr::select(`Route ID`=route_id, 
                `Route Name`=route_long,
                `LEP Households`=total_lep) %>% 
  st_drop_geometry()

kable(manbx_bus_routes_with_lep_kable, 
      caption = paste("Top ", nrow(manbx_bus_routes_with_lep_kable), " bus routes by LEP Household Passbys, MAN/BX")) %>%
  kable_styling(bootstrap_options = "striped")


```

\  

### Applications Needed to Reach a Target Enrollment

Though data has not been collected over sufficient years to make an authoritative prediction about the number of applications needed to recruit to a target enrollment, a rough estimate can be derived from the applications to registrations rate in the years with recorded registrations. 

The below table outlines the current application to registration performance:
  
  ```{r app_to_reg}

recruitment_data %>% 
  st_drop_geometry() %>% 
  count(year, registered) %>% 
  complete(year, registered, fill = list(n = 0)) %>% 
  spread(registered, n) %>% 
  transmute(Year = year,
            `Not Registered` = `FALSE`,
            `Registered` = `TRUE`,
            `Total Applications` = `Not Registered` + `Registered`) %>% 
  kable(., caption = "Total Applications and Registrations, 2017-2019 School Years") %>% 
  kable_styling(bootstrap_options = "striped")



```


Here, the below outlines the estimated applications that might be necessary to hit a target recruitment. This could be used as a hypothesis to test in coming years as further data comes in. A better approach could be looking at application to registration performance in top recruiting neighborhoods to see how they perform over time if recruitment efforts are focused in these neighborhoods over others, or to test whether increased or diversified efforts in underperforming neighborhoods may result in better performance in those areas. 

```{r applications_needed_to_reach_target}

recruitment_data %>% 
  filter(year %in% c(2017, 2018)) %>% 
  st_drop_geometry() %>% 
  count(registered) %>% 
  spread(registered, n) %>% 
  transmute(Applied = `TRUE` + `FALSE`,
            Registered = `TRUE`,
            `Registration Rate` = Registered / Applied,
            `To Reach 1200` = 1200 %/% `Registration Rate`) %>% 
  kable(., caption = "Applications needed to reach target enrollment, based on 2017-2018 recruitment performance") %>% 
  kable_styling(bootstrap_options = "striped")

```