---
title: "HUM II Recruitment"
author: "Ryder Cobean"
output: html_document
---


This report attempts to answer a number of questions for Humanities II Charter School in The Bronx, New York City.





## Special Groups Recruitment
#### Question: If I want to to target and recruit ells, where would the most sensible places be to work on targeted recruitment?

This report reviewed the data from NYC's Department of Education on enrollment statistics in `r length(unique(districts_demo_snapshot_1718$school_dist))` districts, focusing on those in Manhattan and The Bronx. 



```{r message=FALSE, cache=TRUE, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# source("datasets.R", print.eval = F)

tmap_mode("plot")


plot1 <- tm_shape(nyc_area_zips, bbox = c(manhattan_sf, bronx_sf)) +
  tm_fill(col = "grey90") + 
  tm_shape(districts_demo_snapshot_1718, point.per = "feature") + 
  tm_borders() + 
  tm_fill(col = "pct_ell", title = "% ELL Enrollment, 2017-18", palette = "BuGn") +
  tm_text("school_dist") +
  tm_layout(legend.position = c("left", "top"),
            title = "Percentage of ELL enrollment by district, 2017-18",
            title.bg.color = "white", title.bg.alpha = 0.6)




# Plot schools in 95th percentile for ELL enrollment in Manhattan and The Bronx:
high_ell_schools_man_bx <- schools_demo_snapshot_1718 %>% 
  filter(pct_ell > quantile(.$pct_ell, 0.95)) %>% 
  filter(boro %in% c("X", "M")) %>%
  .[, c("dbn", "school_name", "total_enrollment", "pct_ell")]

plot2 <- tm_shape(nyc_area_zips, bbox = c(manhattan_sf, bronx_sf)) + 
  tm_fill(col = "grey90") +
  tm_shape(nyc_sds) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(high_ell_schools_man_bx) + 
  tm_symbols(col = "red",
             size = "pct_ell",
             title.size = "% ELL enrollment",
             perceptual = T,
             sizes.legend = c(40,60,80,100), legend.size.is.portrait = F) + 
  tm_layout(legend.position = c("right", "bottom"),
            legend.bg.color = "white", legend.bg.alpha = 1, legend.frame = "black",
            title = paste("95th-percentile ELL-enrolling schools, Manhattan/Bronx"),
            title.position = c("center", "top"),
            title.bg.color = "white", title.bg.alpha = 0.6)


tmap_arrange(plot1, plot2, ncol = 2, asp = 1)



```

### Schools in Manhattan and the Bronx:
#### 95th percentile for ELL enrollment

Principal Neagley has asked what might be strategies for targeted advertisement to reach out to neighborhoods 

```{r echo=FALSE, message=FALSE}
tmap_mode("view")
tm_shape(nyc_sds) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(high_ell_schools_man_bx, bbox = high_ell_schools_man_bx) + 
  tm_dots(col = "red", 
             size = "pct_ell") + 
  tm_view(bbox = c(manhattan_sf, bronx_sf),
          set.zoom.limits = c(11,15),
          dot.size.fixed = F)
```

